### 2019-03-22 2_metg.tex

To calculate METG, this data is replotted along axes of efficiency
(i.e., as a percentage of the peak FLOPS achieved) and task
granularity (i.e., $\text{wall time} \times \text{number of
  cores}/\text{number of tasks}$). Note that a \emph{task} is defined
broadly to be any continuously-executing unit of work in the
application, thus it makes sense to discuss tasks even in systems
without an explicit notion of tasking, such as in MPI codes are written in
a bulk synchronous style.

Figure~\ref{fig:efficiency-mpi}
shows the replotted data from above. In this chart, we can see that
efficiency starts near 100\%, and that initially task granularity
drops quickly with a minimal change in efficiency. However as task
granularity shrinks, efficiency begins to drop as well, leading
eventually to a plateau in task granularity. Intuitively, this plateau
represents the fundamental baseline overhead of the system, the point
below which the system simply cannot execute tasks efficiently.

>>>>

To calculate METG, this data is replotted along axes of efficiency
(i.e., as a percentage of the peak FLOPS achieved) and task
granularity (i.e., $\text{wall time} \times \text{number of
  cores}/\text{number of tasks}$). Note that a \emph{task} is defined
broadly to be any continuously-executing unit of work in the
application, thus it makes sense to discuss tasks even in systems
without an explicit notion of tasking, such as in MPI codes that are written in
a bulk synchronous style.

Figure~\ref{fig:efficiency-mpi}
shows the replotted data from above. In this chart, we can see that
efficiency starts near 100\%, and that initially task granularity
drops quickly with minimal change in efficiency. However as task
granularity shrinks, efficiency begins to drop as well, leading
eventually to a plateau in task granularity. Intuitively, this plateau
represents the fundamental baseline overhead of the system, the point
below which the system simply cannot execute tasks efficiently.

### 2019-03-22 3_task_bench.tex

For simplicity but without loss of generality, the iteration space in
Task Bench is constrained to be 2-dimensional, with time flowing along
the vertical axis and the width of the task graph along the
horizontal. Tasks can depend only on tasks from the immediately
preceding time step. Many problems can be described by a task graph of
constant width, in which case the iteration space is rectangular, but
Task Bench is capable of handling non-rectangular iteration spaces,
such as required by sweeps and tree-based dependence
patterns. Figure~\ref{fig:task-graphs} shows a number of sample task
graphs that can be implemented with Task Bench.

@@

Despite its generality, Task Bench is easy to implement, making it
tractable to develop a suite of high-quality implementations. Most
Task Bench implementations are several hundred lines long, with the
shortest being 88 lines and the longest being 1500 lines of executable
code. Much of this is possible because the core aspects of generating
task graphs and enumerating dependencies are encapsulated in a core
library that is shared among all the Task Bench implementations. The
core library also includes implementations of the kernels, ensuring
that the kernels are identical in all cases, eliminating a potential
source of performance disparity which is common in mini-apps
implementations. Finally, the Task Bench core library manages the
parsing of input parameters and the display out output results,
ensuring that all implementations behave in a uniform manner, and can
be scripted consistently.

The task implementation in the Task Bench core library is fully
validating. Because the task graph configuration is explicitly
represented (though unmaterialized) in Task Bench, this representation
can be queried to determine exactly what dependencies a task should
expect. The output of every task in Task Bench is uniquely generated,
and all inputs are verified. An assertion is thrown if validation
fails. This ensures that every execution of Task Bench, if it
completes successfully, is correct. This validation provides high
assurance that all Task Bench implementations are correct, and
safeguards against corner cases that may not have been adequately
tested during development. An evaluation of the performance impact of
this validation showed it to be less than 3\% at the smallest task
granularities in any Task Bench implementation, with a negligible
effect on overall results.

>>>>

For simplicity but without loss of generality, the iteration space in
Task Bench is constrained to be 2-dimensional, with time flowing along
the vertical axis and parallel tasks along the
horizontal. Tasks can depend only on tasks from the immediately
preceding time step. Many problems can be described by a task graph of
constant width, in which case the iteration space is rectangular, but
Task Bench is capable of handling non-rectangular iteration spaces,
such as required by sweeps and tree-based dependence
patterns. Figure~\ref{fig:task-graphs} shows a number of sample task
graphs that can be implemented with Task Bench.

@@

Despite its generality, Task Bench is easy to implement, making it
tractable to develop a suite of high-quality implementations. Most
Task Bench implementations are several hundred lines long, with the
shortest being 88 lines and the longest being 1500 lines of executable
code. Much of this is possible because the core aspects of generating
task graphs and enumerating dependencies are encapsulated in a core
library that is shared among all the Task Bench implementations. The
core library also includes implementations of the kernels, ensuring
that the kernels are identical in all cases, eliminating a potential
source of performance disparity which can be a pitfall for mini-apps
implementations. Finally, the Task Bench core library manages the
parsing of input parameters and the display out output results,
ensuring that all implementations behave in a uniform manner, and can
be scripted consistently.

The task implementation in the Task Bench core library is fully
validating. Because the task graph configuration is explicitly
represented (though unmaterialized) in Task Bench, this representation
can be queried to determine exactly what dependencies a task should
expect. The output of every task in Task Bench is uniquely generated,
and all inputs are verified. An assertion is thrown if validation
fails. This ensures that every execution of Task Bench, if it
completes successfully, is correct. This validation provides high
assurance that all Task Bench implementations are correct, and
safeguards against corner cases that may only be exercised at
scale. An evaluation of the performance impact of
this validation showed it to be less than 3\% at the smallest task
granularities in any Task Bench implementation, with a negligible
effect on overall results.

### 2019-03-22 4_implementation.tex

One challenge in implementing Task Bench for such a wide variety of
systems is that the capabilities of the systems vary considerably. For
example, some systems are \emph{implicitly parallel}, and provide some
form of parallelism discovery from sequential programs, whereas others
are \emph{explicitly parallel} and require users to specify the
parallelism in the program. As a rule of thumb, we aim for the Task
Bench implementations to be representative of how actual applications
would be written in the respective systems. Places, where the Task
Bench implementations may be unidiomatic, are noted below.

@@

Spark's programming model is based on \emph{functional parallelism}, a
generalization of the MapReduce programming
model~\cite{MapReduce04}. The core computational abstraction in Spark
are operators such as map, reduce, filter, join, etc. which by virtue
of being pure functions are trivially known to be
parallelizable. Functions in Spark operate on \emph{resilient
  distributed datasets} (RDDs), which are globally-visible,
dynamically single-assignment data structures that cached in memory
and written to disk for durability. One of Spark's primary
optimizations therefore is the caching of RDDs to avoid unnecessary
disk traffic, resulting in substantial speedups over MapReduce.

Though Spark has tasks internally, these are not exposed to the user,
so one of the challenges in developing a Task Bench implementation is
mapping the task graph to a set of operators which will result in the
desired performance behavior. We use a combination of
\lstinline[language=Scala]{flatMap} and
\lstinline[language=Scala]{groupByKey} operations to generate task
dependencies, and then \lstinline[language=Scala]{map} to execute a
series of tasks. We use an explicit hash partitioner to ensure that
Spark does not attempt to group multiple Task Bench tasks into a
single Spark task.

We performed an extensive set of experiments to verify that no
extraneous factors interfered with our Spark measurements. We
completely disabled logging, ensured that no RDDs are actually written
to disk, confirmed that there is no measurable overhead due to JNI
calls from Java to C, and cross-checked our results with known cases
that hit optimal task throughput in Spark, among other things. Our
results are discussed in detail in Section~\ref{sec:evaluation}.

@@

X10~\cite{X1005} is an explicitly parallel programming language for
place-based programming. The core features of X10 are \emph{places}
which represent distributed memories, a PGAS model where references to
remote objects can be held (but can only be deferenced on the local
place), asynchronous tasks, and a place-changing construct to move the
execution of a task to a remote place. X10 also supports a variety of
synchronization primitives.

Our Task Bench implementation uses
\lstinline[language=X10]{Rail.asyncCopy} for efficient data movement
between places, along with place-changing and atomic integers for
synchronization. We use the C++ backend of X10, which uses MPI
internally for portability to a variety of networks, and which
conveniently integrates well with the Task Bench API.

>>>>

One challenge in implementing Task Bench for such a wide variety of
systems is that the capabilities of the systems vary considerably. For
example, some systems are \emph{implicitly parallel}, and provide some
form of parallelism discovery from sequential programs, whereas others
are \emph{explicitly parallel} and require users to specify the
parallelism in the program. As a rule of thumb, we aim for the Task
Bench implementations to be representative of how actual applications
would be written in the respective systems. Places where the Task
Bench implementations may be unidiomatic are noted below.

@@

Spark's programming model is based on \emph{functional parallelism}, a
generalization of the MapReduce programming
model~\cite{MapReduce04}. The core computational abstraction in Spark
are operators such as map, reduce, filter, join, etc. which by virtue
of being pure functions are trivially known to be
parallelizable. Functions in Spark operate on \emph{resilient
  distributed datasets} (RDDs), which are globally-visible,
dynamically single-assignment data structures that are cached in memory
and written to disk for durability. One of Spark's primary
optimizations therefore is the caching of RDDs to avoid unnecessary
disk traffic, resulting in substantial speedups over MapReduce.

Though Spark has tasks internally, these are not exposed to the user,
so one of the challenges in developing a Task Bench implementation is
mapping the task graph to a set of operators which will result in the
desired performance behavior. We use a combination of
\lstinline[language=Scala]{flatMap} and
\lstinline[language=Scala]{groupByKey} operations to generate task
dependencies, and then \lstinline[language=Scala]{mapPartitions} to execute a
series of tasks. We use an explicit hash partitioner to ensure that
Spark does not attempt to group multiple Task Bench tasks into a
single Spark task.

We performed an extensive set of experiments to verify that no
extraneous factors interfered with our Spark measurements. We
completely disabled logging, ensured that no RDDs are actually written
to disk, confirmed that there is no measurable overhead due to JNI
calls from Java to C, and cross-checked our results with known cases
that hit optimal task throughput in Spark, among other things.

@@

X10~\cite{X1005} is an explicitly parallel programming language for
place-based programming. The core features of X10 are \emph{places}
which represent distributed memories, a PGAS model where references to
remote objects can be held (but can only be dereferenced on the local
place), asynchronous tasks, and a place-changing construct to move the
execution of a task to a remote place. X10 also supports a variety of
synchronization primitives.

Our Task Bench implementation uses
\lstinline[language=X10]{Rail.asyncCopy} for efficient data movement
between places, along with place-changing and atomic integers for
synchronization. We use the native backend of X10, which compiles to
C++.

### 2019-03-22 5_evaluation.tex
More related work.
Run spell check.

In theory, any system should be able to hit peak efficiency as long as
the kernels are well-written and of sufficiently large granularity. In
practice, achieving peak performance is tricky, subject to a number of
pitfalls, and unfortunately, frequently skipped in evaluations of
runtime system overhead. However, verifying that peak performance is
achieved is important to ensuring that such evaluation of overhead and
efficiency are well-grounded and avoid common mistakes in
configuration.

Figure~\ref{fig:flops} shows the FLOPS achieved with a compute-bound
kernel with varying problem sizes. This is the full version of
Figure~\ref{fig:flops-mpi}. In the best case, we measure peak FLOPS of
$1.26 \times 10^{12}$, which compares favorably with the officially
reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
of measuring efficiency we use our emperically determined number as
the baseline for 100\% efficiency.

The vast majority of systems achieve at or near peak FLOPS. A number
of systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOPS compared
to systems which use all cores for application tasks. Some of the
slower systems struggle to achieve peak FLOPS, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes it prohibitively
expensive to run larger problem sizes. For example, the Spark job in
this configuration ran for over 5 hours.

Figure~\ref{fig:efficiency} shows the efficiency (as a percentage of
peak FLOPS achieved) plotted against task granularity. This plot is
used to calculate METG, as described in Section~\ref{sec:metg}. The
red, dashed line shows the point where 50\% efficiency is achieved.

In most cases, task granularity has plateaued prior to this point,
though a number of systems continue to improve as efficiency continues
to drop. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold compared to
measuring task scheduling throughput in the presence of empty tasks
(effectively the equivalent of METG(0\%)). Empty tasks permit runtime
systems to make unrealistic tradeoffs which are not applicable to real
application workloads where some degree of efficiency is necessarily a
goal. Reporting numbers with empty tasks therefore gives unfair
advantage to systems that make such unrealistic tradeoffs.

@@

\subsection{Scalability}

>>>>

In theory, any system should be able to hit peak efficiency as long as
the kernels are well-written and of sufficiently large granularity. In
practice, achieving peak performance is tricky, subject to a number of
pitfalls, and unfortunately, frequently skipped in evaluations of
runtime system overhead. However, verifying that peak performance is
achieved is important to ensuring that such evaluations of overhead and
efficiency are well-grounded and avoid common pitfalls in
configuration.

Figure~\ref{fig:flops} shows the FLOPS achieved with a compute-bound
kernel with varying problem sizes. This is the full version of
Figure~\ref{fig:flops-mpi}. In the best case, we measure peak FLOPS of
$1.26 \times 10^{12}$, which compares favorably with the officially
reported number of $1.2 \times 10^{12}$ \cite{Cori}. For the purposes
of measuring efficiency we use our empirically determined number as
the baseline for 100\% efficiency.

The vast majority of systems achieve at or near peak FLOPS. Some
systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOPS compared
to systems which share all cores between the application and runtime. Some of the
slower systems struggle to achieve peak FLOPS, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes it prohibitively
expensive to run larger problem sizes. For example, the Spark job in
this configuration ran for over 6 hours.

Figure~\ref{fig:efficiency} shows efficiency (as a percentage of
peak FLOPS achieved) plotted against task granularity. This plot is
used to calculate METG, as described in Section~\ref{sec:metg}. The
red, dashed line shows the point where 50\% efficiency is achieved.

In most cases, task granularity has plateaued prior to the 50\% efficiency point,
though a number of systems continue to improve as efficiency continues
to drop. Accounting for this effect is one of the main arguments in
favor of METG with a reasonable efficiency threshold compared to
measuring task scheduling throughput in the presence of empty tasks
(effectively the equivalent of METG(0\%)). Empty tasks permit runtime
systems to make unrealistic tradeoffs which are not applicable to real
application workloads where some degree of efficiency is necessarily a
goal. Reporting numbers with empty tasks therefore gives unfair
advantage to systems that make such unrealistic tradeoffs.

@@

\subsection{Scalability}
\label{subsec:scalability}

### 2019-03-22 6_related.tex

A common failure mode for mini-apps is to bite off more complexity
than can be reasonably implemented to a level of quality sufficient
for publication. This can lead to publications for mini-apps omitting
details or performance results for systems beyond the reference
implementation. Follow-up papers, if any, often compare only against a
reference implementation and omit any other implementations that might
exist.

@@

\item
The initial paper on PENNANT~\cite{PENNANT} includes an MPI+OpenMP
implementation; follow-up papers present an implementation in
Regent~\cite{Regent15, ControlReplication17, LegionTracing18}.

@@

The NAS benchmark suite~\cite{NAS91, NAS95} are widely studied with
implementations in OpenMP~\cite{NASOpenMP99}, MPI and
MPI+OpenMP~\cite{NASMPIOpenMP00}, and
Charm++~\cite{NASCharm96}. However the benchmark consider mostly small
kernels for dense array computations, and do not generalize to larger
applications.

>>>>

A common failure mode for mini-apps is to bite off more complexity
than can be reasonably implemented to a level of quality sufficient
for publication. This can lead to publications for mini-apps omitting
implementations or performance results for systems beyond the reference
implementation. Follow-up papers, if any, often compare only against a
reference implementation and omit any other implementations that might
exist.

@@

\item
The initial paper on PENNANT~\cite{PENNANT} includes an MPI/OpenMP/MPI+OpenMP
implementation; follow-up papers present an implementation in
Regent~\cite{Regent15, ControlReplication17, LegionTracing18}.

@@

The NAS benchmark suite~\cite{NAS91, NAS95} are widely studied with
implementations in OpenMP~\cite{NASOpenMP99}, MPI and
MPI+OpenMP~\cite{NASMPIOpenMP00}, and
Charm++~\cite{NASCharm96}. However the benchmarks consist mostly of small
kernels for dense array computations, and do not generalize to larger
applications.

System-specific benchmarks have been used to quantify specific aspects
of system performance, such as MPI communication or collective
latency~\cite{MPPTest99, MPIBench01}. These measurements typically do
not generalize beyond the immediate system they attempt to quantify,
nor in an obvious way to real-world applications.

Limit studies of task scheduling throughput in various runtime systems
often make additional assumptions. A popular assumption is the use of
trivially parallel tasks~\cite{Canary16, Armstrong14}, which as shown
in Section~\ref{subsec:number-of-dependencies} underestimates (often
substantially) the cost of scheduling a task, and can also impact scalability.
